SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization

### 2014-03-19 paper/paper.tex
paper updates.

% can be 9pt
\documentclass{sigplanconf}
\usepackage{url}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{framed}
\usepackage{color}
\usepackage{url}
\usepackage{alltt}
\usepackage{tabularx}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{paralist}

\usepackage{pgf}

@@

\newenvironment{code}{\begin{alltt}}{\end{alltt}}
%% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &

@@

We generalize the idea of ignoring primitive types by introducing the concept of
\emph{opaque types}.  If the reduction algorithm encounters a opaque type, it is
ignored.  Opaque types include the primitive types mentioned above, but in the
user can declare any substructure in a data type to be a opaque type by
providing custom instances, .  Doing so effectively treats values from that type
as ``black boxes'', making SmartCheck more efficient if the user knows that some
portion of the structure cannot be shrunk or is irrelevant to the property.

@@

\subsection{Natural Language Processing}
In 2012, a question was posted on the programming message board
Stack Overflow asking about how to shrink large nested data
types.\footnote{\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}}
The poster writes:
\begin{quote}
$\ldots$ I tend to get an incomprehensible page full of output. $\ldots$
  Implementing the shrink function for each of my types seems to help a little,
  but not as much as I'd like.  $\ldots$ If I try to tune my shrink
  implementations, I also find that QC starts taking a very long time.
\end{quote}

@@

``Retro-fitting'' SmartCheck to GenI, which is a large software system, required
some boilerplate work.  Fortunately, because the program already uses
QuickCheck, \ttp{arbitrary} instances were already defined, but we had to hunt
down and define default instances for a large number of definitions (indeed, the
job could have largely been done with Sed and Awk).

@@

@@

\section*{Acknowledgments}
I thank Andy Gill, Joe Hurd, John Launchbury, Simon Winwood, and especially John
Hughes for reading an earlier draft of this paper.  Their deep insights and bug
catches dramatically improved both the form and content.  This paper was
submitted (and rejected) from ICFP'13; the anonymous reviewers' suggestions
improved the presentation and especially the implementation.

>>>>

% can be 9pt
\documentclass{sigplanconf}
\usepackage{url}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{framed}
\usepackage{color}
\usepackage{url}
\usepackage{alltt}
\usepackage{tabularx}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{paralist}
\usepackage{multirow}
\usepackage{pgf}

@@

\newenvironment{code}{\begin{alltt}\footnotesize}{\end{alltt}}
%% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &

@@

We generalize the idea of ignoring primitive types by introducing the concept of
\emph{opaque types}.  If the reduction algorithm encounters a opaque type, it is
ignored.  Opaque types include the primitive types mentioned above, but in the
user can declare any substructure in a data type to be a opaque type by
providing custom instances.  Doing so effectively treats values from that type
as ``black boxes'', making SmartCheck more efficient if the user knows that some
portion of the structure cannot be shrunk or is irrelevant to the property.

Opaque types can be conditional.  For example, the user may want lists to be
shrunk in general, unless the element of the list are opaque themselves.  Such a
definition is possible.

Opaque types are defined by providing the method for \ttp{opaque} in the
\ttp{SubTypes} type class.

@@

\subsection{Natural Language Processing} \label{sec:nlp}
In 2012, a question was posted on the programming message board
Stack Overflow asking about how to shrink large nested data
types.\footnote{\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}}
The poster writes:
\begin{quote}
$\ldots$ I tend to get an incomprehensible page full of output. $\ldots$
  Implementing the shrink function for each of my types seems to help a little,
  but not as much as I'd like.  $\ldots$ If I try to tune my shrink
  implementations, I also find that QC starts taking a very long time.
\end{quote}

@@

``Retro-fitting'' SmartCheck to GenI, which is a large software system, required
some boilerplate work.  Fortunately, because the program already uses
QuickCheck, \ttp{arbitrary} instances were already defined, but we had to hunt
down the data types used to write \ttp{instance Subtypes D}, for each data type
\ttp{D} (indeed, the job could have largely been done with Sed and Awk).

@@

\subsection{Benchmarks}

No set of testing benchmarks exists over which to compare different test-case
generation and minimization approaches.  Therefore, we have collected a small
number of benchmarks, in addition to the more involved case-studies described
earlier in this section.

The benchmarks presented, in addition to the motivating example presented in
Section~\ref{sec:example}, are

\begin{itemize}
  \item  \emph{Reverse}, with the false property
\begin{code}
prop_rev :: [a] -> Bool
prop_rev ls = ls == reverse ls
\end{code}
\noindent
(the example appears in the original QuickCheck documentation);

  \item \emph{Div0}, a division-by-zero property for a simple calculator
    language (introduced in Section~\ref{sec:subval});

  \item \emph{Heap}, an example from the QuickCheck test suite, in which an
    incorrect ``to sorted list'' function is checked.
\end{itemize}

\noindent
All benchmarks can be found at
\url{https://github.com/leepike/SmartCheck/tree/master/regression}.  Benchmarks
are against QuickCheck~2.7.   The benchmarks only compare the size of shrunk
values and the time required to generate them using standard arguments to
QuickCheck and SmartCheck.  We do not benchmark the generalization algorithms,
since there is no analog in QuickCheck  or other testing frameworks.

\todo{include std dev.}
\todo{fill in benchmarks!}

\begin{table}[ht]
\footnotesize
  \begin{center}
    \begin{tabular}{|r|r||c|c|}
\hline
\multicolumn{2}{|c||}{} & QuickCheck & SmartCheck \\
\hline \hline

\multirow{4}{*}{Reverse}
  & Mean       & 0.08s, 69  & 0.53s, 35  \\
  \cline{2-4}
  & Median     & 0.07s, 69  & 0.17s, 33  \\
  \cline{2-4}
  & Std. dev.  &            &            \\
  \cline{2-4}
  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
  \hline

\multirow{4}{*}{Div0}
  & Mean       & 0.08s, 69  & 0.53s, 35  \\
  \cline{2-4}
  & Median     & 0.07s, 69  & 0.17s, 33  \\
  \cline{2-4}
  & Std. dev.  &            &            \\
  \cline{2-4}
  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
  \hline

\multirow{4}{*}{Heap}
  & Mean       & 0.08s, 69  & 0.53s, 35  \\
  \cline{2-4}
  & Median     & 0.07s, 69  & 0.17s, 33  \\
  \cline{2-4}
  & Std. dev.  &            &            \\
  \cline{2-4}
  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
  \hline
    \end{tabular}

  \end{center}
  \caption{Summarizing data for the graphs in Figure~\ref{fig:graphs}. Entries
    contain execution time (in seconds) and counterexample sizes (counting
    constructors).}
  \label{table:results}
\end{table}

The \emph{Reverse} benchmark essentially provides a lower-bound on the benefit
of using SmartCheck to discover small counterexamples, since a list of length
two falsifies the property (SmartCheck is still useful in generalizing the
counterexample, however!).  Perhaps surprisingly, SmartCheck still slightly
outperforms QuickCheck.  The other two benchmarks are slightly larger examples,
where SmartCheck's benefits are more pronounced.  But again, we emphasize the
most pronounced benefit comes in much larger examples, like described in
Section~\ref{sec:nlp}.

@@

\section*{Acknowledgments}
I thank Andy Gill, Joe Hurd, John Launchbury, Simon Winwood, and especially John
Hughes for reading an earlier draft of this paper.  Their deep insights and bug
catches dramatically improved both the form and content.  This paper was and
rejected from ICFP'13; the anonymous reviewers' suggestions improved the
presentation additional benchmarking.
